---
title: "FBGA Analysis"
author: "EPatton"
date: "2023-01-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(sf)
library(mapview)
library(RColorBrewer)
library(readr)
library(ggplot2)
```

# Project Goals

## Load and clean data for FBGA datasets

### Determine and capture missing data numbers
### Create daily high temps and low temps for data set
### Create sWBGT, if needed

## Trend Analysis

### Increase(decrease) in Red & Black Flag Days for record
### Mann Kendall for entire period daily temps

## Heat Wave Analysis
Heat waves = Red or Black Flag day (?)
### Number of 3+ days heat waves over each month and year
### Number of 5+ days heat waves over each month and year

#### Load Data Sets
```{r load data}
##load the data set, assign it a name, and save the new name
#KLSF_Raw <- KLSF_Observations
#write.csv(KLSF_Raw, "/Users/erikpatton/Library/Mobile Documents/com~apple~CloudDocs/Duke/Data_Dissertation/Raw Data/KLSF_Raw.csv")

KLSF <- read_csv("/Users/erikpatton/Library/Mobile Documents/com~apple~CloudDocs/Duke/Data_Dissertation/Raw Data/KLSF_Raw.csv")
#view(KLSF)
#str(KLSF)
#summary(KLSF)

##CREATE DATE/TIME COLUMN
KLSF <- KLSF %>%
  mutate('datetime' = make_datetime(year=Year, month = Month, day = Day, hour = Hour..UTC.))%>%
  mutate('date' = make_date(year=Year, month = Month, day = Day))
head(KLSF$datetime)

##CREATE CELCIUS COLUMNS, RENAME COLUMNS TO STANDARIZE
KLSF <- KLSF%>%
  mutate('Temp.C' = ((Temperature..F.-32)*(5/9)))%>%
  mutate('Dewpoint.Temp.C' = (Dewpoint.Temperature..F.-32)*(5/9))%>%
  mutate('Heat.Index.C'=((Heat.Index..F.-32)*(5/9)))%>%
  mutate('WBGT.C'=((Derived.Wet.Bulb.Globe.Temperature..F.-32)*(5/9)))%>%
  rename('Temp.F' = Temperature..F.)%>%
  rename('Dewpoint.Temp.F'=Dewpoint.Temperature..F.)%>%
  rename('Heat.Index.F'=Heat.Index..F.)%>%
  rename("WBGT.F"=Derived.Wet.Bulb.Globe.Temperature..F.)
```

```{r CREATE SINGLE DAY MAX or MIN TEMPS}
KLSF_DAILY_HIGH.C <- KLSF%>%
  group_by(date)%>%
  dplyr::summarise(daily_high = max(Temp.C))%>%
  as.data.frame()

KLSF_DAILY_LOW.C <- KLSF%>%
  group_by(date)%>%
  dplyr::summarise(daily_high = min(Temp.C))%>%
  as.data.frame()
```

#### Create an initial plot to visualize missing data 
```{r}
KLSF_DAILY_HIGH_PLOT.C <- ggplot(KLSF_DAILY_HIGH.C, aes(x = date, y=daily_high))+
  geom_line()+
  geom_smooth(method=lm, col='red')
print(KLSF_DAILY_HIGH_PLOT.C)
```
#### LOAD HEAT CATAGORY TEMPERATURE THRESHOLDS
```{r TEMPERATURE THRESHOLDS}
##Create the heat catagory data frame for reference
Flag_Color <- c("Green","Yellow","Red","Black")
Category_Temp_Min.F <- c(82,85,88,90)
Category_Temp_Max.F <- c(84.999,87.999,89.999,104)
Heat_Categories <- cbind(Flag_Color,Category_Temp_Min.F,Category_Temp_Max.F)
Heat_Categories <- as.data.frame(Heat_Categories)
##recast the vector as an integer to get the mutate code to work
Heat_Categories$Category_Temp_Min.F <- as.integer(Heat_Categories$Category_Temp_Min.F)
Heat_Categories$Category_Temp_Max.F <- as.integer(Heat_Categories$Category_Temp_Max.F)
##Create degree Celsius columns
Heat_Categories <- Heat_Categories%>%
  mutate(Category_Temp_Min.C = (Category_Temp_Min.F - 32)*(5/9))%>%
  mutate(Category_Temp_Max.C = (Category_Temp_Max.F-32)*(5/9))

#view(Heat_Categories)
```

