---
title: "FBGA Analysis"
author: "EPatton"
date: "2023-01-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(sf)
library(mapview)
library(RColorBrewer)
library(readr)
library(ggplot2)
```

# Project Goals

## Load and clean data for FBGA datasets

### Determine and capture missing data numbers (done)
### Create daily high temps and low temps for data set (done)
### Create sWBGT, if needed

## Trend Analysis

### Increase(decrease) in Red & Black Flag Days for record
### Mann Kendall for entire period daily temps

## Heat Wave Analysis
Heat waves = Red or Black Flag day (?)
### Number of 3+ days heat waves over each month and year
### Number of 5+ days heat waves over each month and year

#### Load Data Sets
```{r load data}
##load the data set, assign it a name, and save the new name
#KCSG_Raw <- KCSG_Observations
#write.csv(KCSG_Raw, "/Users/erikpatton/Library/Mobile Documents/com~apple~CloudDocs/Duke/Data_Dissertation/Raw Data/KCSG_Raw.csv")

KCSG <- read_csv("/Users/erikpatton/Library/Mobile Documents/com~apple~CloudDocs/Duke/Data_Dissertation/Raw Data/KCSG_Raw.csv")
#view(KCSG)
#str(KCSG)
#summary(KCSG)

##CREATE DATE/TIME COLUMN
KCSG <- KCSG %>%
  mutate('datetime' = make_datetime(year=Year, month = Month, day = Day, hour = Hour..UTC.))%>%
  mutate('date' = make_date(year=Year, month = Month, day = Day))
head(KCSG$datetime)

##CREATE CELCIUS COLUMNS, RENAME COLUMNS TO STANDARIZE
KCSG <- KCSG%>%
  mutate('Temp.C' = ((Temperature..F.-32)*(5/9)))%>%
  mutate('Dewpoint.Temp.C' = (Dewpoint.Temperature..F.-32)*(5/9))%>%
  mutate('Heat.Index.C'=((Heat.Index..F.-32)*(5/9)))%>%
  mutate('WBGT.C'=((Derived.Wet.Bulb.Globe.Temperature..F.-32)*(5/9)))%>%
  rename('Temp.F' = Temperature..F.)%>%
  rename('Dewpoint.Temp.F'=Dewpoint.Temperature..F.)%>%
  rename('Heat.Index.F'=Heat.Index..F.)%>%
  rename("WBGT.F"=Derived.Wet.Bulb.Globe.Temperature..F.)
```

```{r CREATE SINGLE DAY MAX or MIN TEMPS}
KCSG_DAILY_HIGH.C <- KCSG%>%
  group_by(date)%>%
  dplyr::summarise(daily_high = max(Temp.C))%>%
  as.data.frame()

KCSG_DAILY_LOW.C <- KCSG%>%
  group_by(date)%>%
  dplyr::summarise(daily_high = min(Temp.C))%>%
  as.data.frame()
```

#### Create an initial plot to visualize missing data 
```{r}
KCSG_DAILY_HIGH_PLOT.C <- ggplot(KCSG_DAILY_HIGH.C, aes(x = date, y=daily_high))+
  geom_line()+
  geom_smooth(method=lm, col='red')
print(KCSG_DAILY_HIGH_PLOT.C)
```
#### LOAD HEAT CATAGORY TEMPERATURE THRESHOLDS
```{r TEMPERATURE THRESHOLDS}
##Create the heat catagory data frame for reference
Flag_Color <- c("Green","Yellow","Red","Black")
Category_Temp_Min.F <- c(82,85,88,90)
Category_Temp_Max.F <- c(84.999,87.999,89.999,104)
Heat_Categories <- cbind(Flag_Color,Category_Temp_Min.F,Category_Temp_Max.F)
Heat_Categories <- as.data.frame(Heat_Categories)
##recast the vector as an integer to get the mutate code to work
Heat_Categories$Category_Temp_Min.F <- as.integer(Heat_Categories$Category_Temp_Min.F)
Heat_Categories$Category_Temp_Max.F <- as.integer(Heat_Categories$Category_Temp_Max.F)
##Create degree Celsius columns
Heat_Categories <- Heat_Categories%>%
  mutate(Category_Temp_Min.C = (Category_Temp_Min.F - 32)*(5/9))%>%
  mutate(Category_Temp_Max.C = (Category_Temp_Max.F-32)*(5/9))

#view(Heat_Categories)
```

#### Start Red and Black Flag Day Analysis
```{r}
##Filter for all days in 'RED' or 'BLACK' heat category
FBGA.KCSG_WBGT_Filtered <- KCSG%>%
  filter(KCSG$WBGT.F >= Heat_Categories$Category_Temp_Min.F[3])%>%
  drop_na(WBGT.F)
#view(FBGA_WBGT_Filtered)
summary(FBGA.KCSG_WBGT_Filtered$WBGT.F)

##visualize the days above RED and BLACK heat cat
KCSG_REDorBLACK_plot <- ggplot(FBGA.KCSG_WBGT_Filtered, aes(x = date, y=WBGT.F))+
  geom_line()+
  geom_point()
print(KCSG_REDorBLACK_plot)

```
#### Group by year, count and  plot days above RED and above BLACK
```{r}
##group the data by year to do a trend analysis over the years - from 1973 to 2022
##returns number of hours per year at or above RED
##Filter for hours at RED or above
FBGA.KCSG_WBGT_REDorABOVE_Hourly_byYear <- KCSG%>%
  filter(date>='1970-01-01')%>%
  subset(WBGT.F>=Heat_Categories$Category_Temp_Min.F[3])%>%
  group_by(Year)%>%
  count(Year)%>%
  rename(Hours_Above_Red = n)
view(FBGA.KCSG_WBGT_REDorABOVE_Hourly_byYear)
##Plots hours Red or Above
KCSG_Hours_REDorABOVE_peryear_plot <- ggplot(FBGA.KCSG_WBGT_REDorABOVE_Hourly_byYear, aes(x = Year, y=Hours_Above_Red))+
  geom_line()+
  geom_point()
print(KCSG_Hours_REDorABOVE_peryear_plot)

##Filter for hours BLACK
FBGA.KCSG_Hours_BLACK_byYear <- KCSG%>%
  filter(date>='1970-01-01')%>%
  subset(WBGT.F>=Heat_Categories$Category_Temp_Min.F[4])%>%
  group_by(Year)%>%
  count(Year)%>%
  rename(Hours_Above_Black = n)
#view(FBGA.KCSG_Hours_BLACK_byYear)
##Plot hours in Black conditions
KCSG_Hours_Black_peryear_plot <- ggplot(FBGA.KCSG_Hours_BLACK_byYear, aes(x = Year, y=Hours_Above_Black))+
  geom_line()+
  geom_point()
print(KCSG_Hours_Black_peryear_plot)
```

